---
title: "working with igraph package"
author: "Maria Prokofieva"
date: "18/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(statnet)
library(tidygraph)
library(ggraph)
```

### Working with `igraph`

2.1 Creating a network object in `igraph`

### Creating network objects in `igraph``

Let's create an *undirected* graph with 3 edges: 1–>2, 2–>3, 3–>1.

```{r}
g1 <- graph(edges=c(1,2, 2,3, 3, 1), n=3, directed=F)
class(g1)

g1
plot(g1) # simple plot of the network
```

Let's do more!

```{r}
g2 <- graph( edges=c(1,2, 2,3, 3, 1), n=10, directed=F)
class(g2)
g2
plot(g2) # how does it look like?

```

Let's move on to some specific graphs and graph models

**Empty graph**: `graph.empty` or `make_empty_graph`
```{r}
g<-graph.empty(n=10, directed=TRUE)

#g<-make_empty_graph(n=10, directed=TRUE)
plot(g)
```

**Complete graph**: each pair of vertices has an edge connecting them, we use `LOOPS = FALSE` to not add self edges

We can use `graph.full` OR `make_full_graph`

```{r}
g<-graph.full(n=10, directed = FALSE, loops = FALSE)
plot(g)
```

**Stars**
```{r}
g<-graph.star(n=10, mode="out") #or make_star()
plot(g)

g<-graph.star(n=10, mode="in")
plot(g)
```

**Rings**
```{r}
g<-graph.ring(n=10) #or make_ring()
```
**Task**: 

- create a star graph with 7 vertices pointed to the center

- create a full directed graph with 12 vertices and no loops

`igraph` can also create some notable historical graphs, such as [Zachary carate club](https://en.wikipedia.org/wiki/Zachary%27s_karate_club)

```{r}
 zach <- graph("Zachary") # the Zachary carate club
zach
 plot(zach, vertex.size=10, vertex.label=NA)
```

The description of the igraph object typically start with up to four letters:

`D` or `U`, for a directed or undirected graph

`N` for a named graph (where nodes have a name attribute)

`W` for a weighted graph (where edges have a weight attribute)

`B` for a bipartite (two-mode) graph (where nodes have a type attribute)

The two numbers that follow refer to the number of nodes and edges in the graph. 

The description also lists graph, node & edge attributes, e.g.:

`(g/c)` - graph-level character attribute

`(v/c)` - vertex-level character attribute

`(e/n)` - edge-level numeric attribute

Now let's create more "complicated" graphs with the given list of edges

```{r}
edges <- c(1,2, 3,2, 2,4)
g<-graph(edges, n=max(edges), directed=TRUE)
plot(g)
```

Can we do a **named graph**? Sure we can!

When you use vertex names in your edge list, you don't need the number of nodes

```{r}
g3 <- graph( c("expresso", "milk", "milk", "chocolate", "cream", "chocolate", "expresso", "cream")) # named vertices 
plot(g3)
```
For smaller graphs we can use

- ` - `for undirected tie, 

- `+-` for pointing left or 

- `-+` for directed ties  & right, 

- `++` for a symmetric tie, and 

- “:” for sets of vertices.

```{r}

plot(graph_from_literal("dogs"---"cats", "cats"---"mice")) # the number of dashes doesn't matter

```
**Task**

- create a named graph `family` with the nodes=your family members/relates/friends - try to keep it under 20!

Use the directed graph to show your family tree and connections

------------------

### Graph structure information

To access vertices and edges:

```{r}
E(g3) # The edges of the object

V(g3) # The vertices of the object
```
To view the network in a matrix form

```{r}
g3[]
```

`vcount` and `ecount` return the number of vertex / number of edges

```{r}
edges <- c(1,2, 3,2, 1,4)
g<-graph(edges, n=max(edges), directed=TRUE)
vcount(g)
ecount(g)
```

`neighbors` returns the number of neighbors for the vertex. A vertex is a **neighbor** of another one, if they are incident to the same edge, i.e.  the two vertices are adjacent

```{r}

neighbors(g, V(g)[1])

zachary_graph <- make_graph("Zachary")
plot(zachary_graph)
n17 <- neighbors(zachary_graph, 17)
n7 <- neighbors(zachary_graph, 7)
intersection(n17, n7)

```

`is_directed`: directed/undirected  type of graph  

`are_adjacent`:is there a link between vertices v1 and v2 in graph 

```{r}
is_directed(g)

are_adjacent(zachary_graph, n17, n7)

```

`get.edgelist`: returns the list of edges in a graph. 



```{r}
get.edgelist(g)
```

`get.adjacency`: to get the adjacency matrix of a graph.

```{r}
get.adjacency(zachary_graph)

```

**Task**

In the previous task you created the named graph `family` with the nodes=your family members/relates/friends:

- show the number of verices/edges in your graph 

- display the edgelist and adjacency matrix

----------

### Working with network attributes, edges and verteces

Compared to the matrix, network objects have the ability to store additional **attribute** information about the nodes and ties within the same network object.

Add attributes to the network, vertices, or edges:

```{r}
V(g3)$name # automatically generated when we created the network.

V(g3)$calories <- c("high", "medium", "high", "high")
E(g3)$type <- "yumminess" # Edge attribute, assign "yumminess" to all edges
E(g3)$weight <- 10 # Edge weight, setting all existing edges to 10
```
To examine attributes:
```{r}
edge_attr(g3)
vertex_attr(g3)
graph_attr(g3)
```
We can also set attributes using `set_edge_attr()`, `set_vertex_attr()`

```{r}
g3 <- set_graph_attr(g3, "calories", "yumminess")
g3 <- set_graph_attr(g3, "time of day", "time")

graph_attr_names(g3)
graph_attr(g3, "calories")
graph_attr(g3)
```

Let's delete the attribute 
```{r}
g3 <- delete_graph_attr(g3, "time of day")
graph_attr(g3)
plot(g3, edge.arrow.size=.5, vertex.label.color="red", vertex.label.dist=1.5,
     vertex.color=c("green", "yellow")[1+(V(g3)$calories=="high")] ) 

```

How can we `simplify` the graph using `edge.attr.comb` to combine edge attributes:

Options: 

- `sum`, `mean`, `prod` (=product), `min`, `max`, `first`/`last` (selects the first/last edge's attribute). 

- `ignore` to disregard/drop  the attribute 

```{r}

g3_simplified <- simplify(g3, remove.multiple = T, remove.loops = F, 
                 edge.attr.comb=list(weight="sum", type="ignore") )
plot(g3_simplified, vertex.label.dist=1.5)
g3_simplified
```

**Task**

In the previous task you created the named graph `family` with the nodes=your family members/relates/friends:

- add an attribute to vertices (e.g gender)

- add an attribute to edges specifying the type of relations (e.g. good, love, well_it_is_family)

- remove the last attribute you added to edges (they'd better not know!)
----------

### Modifying the graph

You can change the graph by adding more vertices and/or edges

```{r eval=FALSE}

add_edges(graph, edges, ..., attr=list())
add_vertices(graph, vertices, ..., attr=list())
```

Let's create an empty graph and add edges and vertices to it. We will be adding an attribute `color` to edges and using a new value for `color` for new members.

```{r}
g <- make_empty_graph(n = 7) %>%
  add_edges(c(1,2, 2,3, 3,4, 4,5, 6,7)) %>%
  set_edge_attr("color", value = "green") %>%
  add_edges(c(5,1), color = "red")
E(g)[[]]
plot(g)
```

Let's do the same for vertices

```{r eval=FALSE}
add_vertices(graph, nv, ..., attr = list())
```

where `nv` is the number of vertices to add

```{r}
g <- g %>%
  add_vertices(8, color = "pink") %>%
  add_vertices(9, color = "purple") %>%
  add_edges(c(7,8, 8,1, 9,1, 9,2))
g
V(g)[[]]
plot(g)
  
```
**Task**

In the previous task you created the named graph `family` with the nodes=your family members/relates/friends:

### Graphs import / export

We are going to talk about importing graphs in more details when we are going to  go through case studies. For now, you can use the following functions from `igraph`:

- `read.graph()` to load graph using edgelist

- `graph.adjacency()` from adjacency matrix: you can load your matrix using `read_csv()` (from `tidyverse`) to load the dataset, then convert it to a matrix using `as.matrix()` function and then using this matrix in `graph.adjacency()`

- `graph.data.frame()` to import from a dataframe/table

To export your graph you can use `write_graph()`to save it in different formats.

```{r eval=FALSE}
write_graph(
  graph,
  file,
  format = c("edgelist", "pajek", "ncol", "lgl", "graphml", "dimacs", "gml", "dot",
    "leda")
)
```


**Task**

In the previous task you created the named graph `family` with the nodes=your family members/relates/friends:

Save your graph as a pdf file in your working directory
  -

### Graph visualization

Graph visualization is a HUGE topic and we are going to talk about that in details in the second half of the workshop

-----

**Final tasks to summarize**

Create a ring graph with random number of vertex (between 10 and 15). *HINT* use `sample() function to generate the random number. 

```{r eval=FALSE}
sample(x, size, replace = FALSE, prob = NULL)

```
where  

- `x` is a vector of elements to choose from

- `size` is a non-negative integer giving the number of items to choose.

Use the functions covered earlier to answer the following questions (do not plot!):

1. How many edges and vertices in this graph? 

2. Print neighbors of the 10th vertex, the incident edges of this vertex. 

3. Are the 3rd and the 13th vertices connected? 

4. Are the 2nd and the 5th vertices also connected ? 

5. Now, DO a plot of the graph :)

*Solutions*
```{r}

n <-sample(10:15,1)

g <- graph.ring(n)

vcount(g)

ecount(g)

neighbors(g, 10)

incident(g, 10)


are.connected(g, 3, 10)

are.connected(g, 2, 5)

plot(g, layout = layout.fruchterman.reingold,vertex.label=V(g)$number, edge.arrow.size=0.5)

```
---------
Let's do more:

1. Create an empty undirected graph with 6 vertices. 

2. Add edges between the following nodes: 1 and 3, 1 and 5, 5 and 6,  and 1 and 5. 

3. Plot this graph. 

4. Add the 7th vertex and connect it to that which has the most number of neighbors. 

5. Assign names to all vertices (e.g. use days of weeks) and some random weights to all of edges (see the previous example for random values between 0 and 1). 

6. Print the adjacency matrix with weights.


*Solutions*
```{r}

g <- graph.empty (5, directed = FALSE)

new_edges <- c(1,3, 1,5, 2,5, 4,5)

g <- add_edges(g, new_edges)

plot(g)

g <- add_vertices(g, 1)

g <- add_edges(g, c(6,5))

V(g)$name <- letters[1:vcount(g)]

E(g)$weight <- runif(ecount(g))

get.adjacency(g, attr="weight")

plot(g, layout = layout.fruchterman.reingold,vertex.label=V(g)$number)
```

Create a star graph with 8 edges. Next, assign random numbers between 1 and 50 to vertices and select vertices for which this value is less than 30, set the color of these vertices to green.

*Solutions*
```{r}

g <- graph.star(8)

V(g)$number <- sample(1:50, vcount(g), replace=TRUE)

V(g)$color <- "grey"

V(g)[ number < 30 ]$color <- "green"

plot(g, layout=layout.circle, vertex.color=V(g)$color,

         vertex.label=V(g)$number)

```

Create a complete graph with 5 edges. Next, assign random weights between 0 and 1 to all of edges. Set width of edges to 2 and color to green for those of them that have weight less than 0.5, set the width value to 1, color to red for others.

*Solutions*
```{r}

g <- graph.full(5)

E(g)$weight <- runif(ecount(g))

E(g)$width <- 1

E(g)$color <- "red"

E(g)[ weight < 0.5 ]$width <- 2

E(g)[ weight < 0.5 ]$color <- "green"

plot(g)

```

---------

## Case study - French trains

Link







