---
title: "igraph2 - French trains case study"
author: "Maria Prokofieva"
date: "18/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(igraph)
library(statnet)
library(tidygraph)
library(ggraph)
```

##Case study - French trains

Let's look at using network science in more details using a more practical example.

We are going to use this dataset from `TidyTuesday` which is an awesome project run in the R community.This is a weekly project which means that every week they post a raw dataset, a chart or article related to that dataset, and ask you to explore the data. 

The dataset is "almost" tidy and give an excellent opportunity to practice R skills to wrangle the data and wreck your brains
[Link](https://github.com/rfordatascience/tidytuesday)

This time we will look at one of the earlier datasets published [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-02-26) 

The data comes from the SNCF (National Society of French Railways) open data [portal](https://ressources.data.sncf.com/explore/dataset/regularite-mensuelle-tgv-aqst/information/?sort=nombre_de_trains_annules&q=gare&dataChart=eyJxdWVyaWVzIjpbeyJjaGFydHMiOlt7InR5cGUiOiJzcGxpbmUiLCJmdW5jIjoiQVZHIiwieUF4aXMiOiJleHRlcm5lcyIsInNjaWVudGlmaWNEaXNwbGF5Ijp0cnVlLCJjb2xvciI6IiNhNmQ4NTQifV0sInhBeGlzIjoicGVyaW9kZSIsIm1heHBvaW50cyI6IiIsInRpbWVzY2FsZSI6Im1vbnRoIiwic29ydCI6IiIsImNvbmZpZyI6eyJkYXRhc2V0IjoicmVndWxhcml0ZS1tZW5zdWVsbGUtdGd2LWFxc3QiLCJvcHRpb25zIjp7InNvcnQiOiJub21icmVfZGVfdHJhaW5zX2FubnVsZXMifX19XSwidGltZXNjYWxlIjoiIiwiZGlzcGxheUxlZ2VuZCI6dHJ1ZSwiYWxpZ25Nb250aCI6dHJ1ZX0%3D).

-------

The SNCF is France's national state-owned railway company. Founded in 1938, it operates the country's national rail traffic along with Monaco, including the TGV, France's high-speed rail network. 

This dataset covers 2015-2018 with a lot of different train stations.

It comes in two versions: 

- `full_trains.csv` dataset with it's 27 columns

- `small_trains.csv` dataset with only 13 columns

We will be working primarily with `small_trains.csv` for demonstation purposes.



--------

Let's start by reading the file and creating a network object

```{r}
small_trains <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-26/small_trains.csv")

head(small_trains)
```

Let's create the **node list**:

We need to isolate `distinct` stations from both the “departure_station” and “arrival_station” columns and then join the information from these columns together.

```{r}
small_trains <- small_trains %>%  
  group_by(departure_station) %>%
  mutate(late_depart = mean(num_late_at_departure), 
            late_arrive=mean(num_arriving_late))%>% 
  ungroup()

routes_from <- small_trains %>%
  distinct(departure_station, service) %>%
  rename(label = departure_station)

nodes_attributes<-small_trains %>%
  select(departure_station, late_depart, late_arrive)

head(routes_from)

routes_to <- small_trains %>%
  distinct(arrival_station) %>%
  rename(label = arrival_station)

head(routes_to)
```

And use a full join to include all unique places from both the departure_station and arrival_station.

```{r}
nodes <- full_join(routes_from, routes_to, by = "label")
head(nodes)
```
Let's give unique IDs for each station

```{r}
nodes <- nodes %>% 
  rowid_to_column("id")
head(nodes)
```

Let's create the **edge list**:

```{r}
per_route <- small_trains %>%  
  group_by(departure_station, arrival_station) %>%
  summarise(num_trips = mean(total_num_trips), 
            journey_time_avg=mean(journey_time_avg))%>% 
  ungroup()

head(per_route)
```
But... we still have “departure_station” and “arrival_station” columns contain labels rather than IDs. 

Let's link the IDs that have been assigned in nodes to each station in both the “departure_station” and “arrival_station” columns.

```{r}
edges <- per_route %>% 
  left_join(nodes, by = c("departure_station" = "label")) %>% 
  rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("arrival_station" = "label")) %>% 
  rename(to = id)

head(edges)
```
Let's reorder the columns so that the dataset starts with “from” and “to” (IDs). We also do not need “departure_station” and “arrival_station” columns with the names of the stations.

Let's use `select() function` to include the “from”, “to”, and “num_trips",  "journey_time_avg” columns.

```{r}
edges <- edges%>%
  select(from, to, num_trips,  journey_time_avg)

head(edges)
```

Create network object

```{r}
routes_graph <- graph_from_data_frame(d=edges, vertices=nodes, directed=T)

#write_graph(routes_graph, "routes_graph.graphml", format="graphml")

#routes_network <- network(edges, vertex.attr = nodes, matrix.type = "edgelist", ignore.eval = FALSE)

#routes_network<-read_graph("routes_graph.graphml", format="graphml")

```

`graph.data.frame()`function creates an igraph graph from one or two data frames containing the edge list and edge/vertex attributes.

```{r eval=FALSE}
graph.data.frame(d, directed = TRUE, vertices = NULL)
```

where

`d`: a dataframe containing a symbolic edge list in the first two columns. Additional columns are considered as edge attributes. 

`directed`:	Logical scalar, whether or not to create a directed graph.

`vertices`:	 a dataframe with vertex metadata, or NULL. 

Let's look at `routes_graph` closer

```{r}
class(routes_graph)
routes_graph
```

As we learned earlier we  have easy access to nodes, edges, and their attributes with:
```{r}
E(routes_graph) # The edges of the "routes_graph" object
V(routes_graph) # The vertices of the "routes_graph" object
E(routes_graph)$weight # Edge attribute "weight"
V(routes_graph)$name # Vertex attribute "name"
```

###Basic visualization of network

Now that we have our `igraph network object`, let’s plot it.

```{r}
plot(routes_graph, edge.arrow.size=.4,vertex.label=NA)
```
That doesn’t look very good and needs further work...

We can extract an edge list or a matrix from igraph networks.

```{r}
as_edgelist(routes_graph, names=T)  %>% head()
as_adjacency_matrix(routes_graph)  %>% head()
```

Or data frames describing nodes and edges:

```{r}
as_data_frame(routes_graph, what="edges")   %>% head()
as_data_frame(routes_graph, what="vertices")   %>% head()
```

But back to plotting with igraph

`igraph` allows to use a wide set of parameters for plotting: 

- node options (starting with `vertex.`) and 

- edge options (starting with `edge.`). 

A list of selected options is included below, but you can also see them if you use `?igraph.plotting`.

**NODES**

Option | Use
---|---
vertex.color | Node color
vertex.frame.color | Node border color
vertex.shape  | One of “none”, “circle”, “square”, “csquare”, “rectangle”, “crectangle”, “vrectangle”, “pie”, “raster”, or “sphere”
vertex.size | Size of the node (default is 15)
vertex.size2 | The second size of the node (e.g. for a rectangle)
vertex.label | Character vector used to label the nodes
vertex.label.family | Font family of the label (e.g.“Times”, “Helvetica”)
vertex.label.font | Font: 1 plain, 2 bold, 3, italic, 4 bold italic, 5 symbol
vertex.label.cex | Font size (multiplication factor, device-dependent)
vertex.label.dist | Distance between the label and the vertex
vertex.label.degree | The position of the label in relation to the vertex, where 0 right, “pi” is left, “pi/2” is below, and “-pi/2” is above

-----------

**EDGES**

------------

Option | Use
---|---
edge.color | Edge color
edge.width | Edge width, defaults to 1
edge.arrow.size | Arrow size, defaults to 1
edge.arrow.width | Arrow width, defaults to 1
edge.lty | Line type, could be 0 or “blank”, 1 or “solid”, 2 or “dashed”, 3 or “dotted”, 4 or “dotdash”, 5 or “longdash”, 6 or “twodash”
edge.label | Character vector used to label edges
edge.label.family | Font family of the label (e.g.“Times”, “Helvetica”)
edge.label.font | Font: 1 plain, 2 bold, 3, italic, 4 bold italic, 5 symbol
edge.label.cex | Font size for edge labels
edge.curved | Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
arrow.mode | Vector specifying whether edges should have arrows, possible values: 0 no arrow, 1 back, 2 forward, 3 both

-----

**OTHER**

-----

Option | Use
---|---
margin | Empty space margins around the plot, vector with length 4 frame if TRUE, the plot will be framed
main | If set, adds a title to the plot
sub | If set, adds a subtitle to the plot


Let's start changing the plot:

We can set the node & edge options in two ways - the first one is to specify them in the plot()
function, as we are doing below.

```{r}
# Plot with curved edges (edge.curved=.1) and reduce arrow size:
plot(routes_graph, edge.arrow.size=.4, edge.curved=.1, vertex.label=NA)
```
```{r}
# Set edge color to gray, and the node color to orange.
# Replace the vertex label with the node names stored in "media"
plot(routes_graph, edge.arrow.size=.2, edge.curved=0,
vertex.color="orange", vertex.frame.color="#555555",
vertex.label=V(routes_graph)$name, vertex.label.color="black",
vertex.label.cex=.5)
```
The alternative way is to add attributes to the igraph object. 

For example, we can add color or size network nodes based on location. We can alsochange the width of the edges based on their weight.

FIX
```{r eval=FALSE}
# Generate colors based on media type:
colrs <- c("gray50", "tomato", "gold")
V(net)$color <- colrs[V(net)$media.type]
# Set node size based on audience size:
V(net)$size <- V(net)$audience.size*0.7
# The labels are currently node IDs.
# Setting them to NA will render no labels:
V(net)$label.color <- "black"
V(net)$label <- NA
# Set edge width based on weight:
E(net)$width <- E(net)$weight/6
#change arrow size and edge color:
E(net)$arrow.size <- .2
E(net)$edge.color <- "gray80"
E(net)$width <- 1+E(net)$weight/12
```

we can also add a legend explaining the meaning of the colors we used:

```{r eval=FALSE}
plot(net)
legend(x=-1.5, y=-1.1, c("Newspaper","Television", "Online News"), pch=21,
col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
```

For semantic networks, we may plot only the labels of the nodes:

```{r}
plot(routes_graph, vertex.shape="none", vertex.label=V(routes_graph)$name,
vertex.label.font=2, vertex.label.color="gray40",
vertex.label.cex=.7, edge.color="gray85")
```

#### Network layouts

**Network layouts** are  algorithms that return coordinates for each node in a network.

We can set the layout in the plot function:

`igraph` has a number of built-in layouts, including:

```{r}
# Randomly placed vertices
l <- layout_randomly(routes_graph)
plot(routes_graph, layout=l)
```
```{r}
# Circle layout
l <- layout_in_circle(routes_graph)
plot(routes_graph, layout=l)
```
```{r}
# 3D sphere layout
l <- layout_on_sphere(routes_graph)
plot(routes_graph, layout=l)
39
```

One of the most used force-directed layout algorithms is `Fruchterman-Reingold`.

`Force-directed layouts` is aimed nice-looking graph where edges are similar in length and cross
each other as little as possible:

- They simulate the graph as a physical system. 

- Nodes are electrically charged particles that repulse each other when they get too close. 

- The edges act as springs that attract connected nodes closer together. 

- Thus, nodes are evenly distributed: nodes which share more connections are closer to each
other. 

Such algorithms can get slow, especially if you have a larger graph (e.g. 1000 vertices+)

You can use the `weight` to increase the attraction forces among nodes connected by heavier edges.

```{r}
l <- layout_with_fr(routes_graph)
plot(routes_graph, layout=l)
```
The layout is not deterministic - different runs will result in slightly different
configurations. Saving the layout in l allows us to get the exact same result multiple times, which
can be helpful if you want to plot the time evolution of a graph, or different relationships – and
want nodes to stay in the same place in multiple plots.
par(mfrow=c(2,2), mar=c(0,0,0,0)) # plot four figures - 2 rows, 2 columns
plot(net.bg, layout=layout_with_fr)
plot(net.bg, layout=layout_with_fr)
plot(net.bg, layout=l)
plot(net.bg, layout=l)
40

```{r}
l <- layout_with_drl(routes_graph)

plot(routes_graph, layout=l)

```

------------
#### Measures

**Density**: The proportion of present edges from all possible edges in the network.

```{r}
edge_density(routes_graph)
```

**Reciprocity** The proportion of reciprocated ties (for a directed network) (i.e. probability that the opposite counterpart of a directed edge is also included in the graph)

```{r}
reciprocity(routes_graph)
```

**Transitivity**:

- **global** - ratio of triangles (direction disregarded) to connected triples.

- **local** - ratio of triangles to connected triples *each vertex is part of*. This is calculated for each vertex given in the `vids` argument of  `transitivity()`.

```{r eval=FALSE}
transitivity(routes_graph, type="global") # routes_graph is treated as an undirected network
transitivity(routes_graph, mode="local") # same as above
```


**Diameter**:
A network diameter is the longest geodesic distance = length of the shortest path between two nodes in the network. 

`diameter()` returns the distance vs `get_diameter()` returns the nodes along the first found path of that distance.

```{r}
diameter(routes_graph, directed=F, weights=NA)
diameter(routes_graph, directed=F)
```


