---
title: "Practical Applications of Network Science: Workshop "
subtitle: "⚔<br/>with xaringan"
author: "Dr Maria Prokofieva"
institute: "VU Business School"
date: "2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"

      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center")
```
#Practical Applications of Network Science: Workshop:


theory and practical applications of network science to solve today's challenges
---
#Packages to use

[igraph](https://igraph.org/r/)

[statnet](https://cran.r-project.org/web/packages/statnet/)

[sna](https://cran.r-project.org/web/packages/sna/)

```{r}

#load libraries
library (igraph)
library (statnet)
library (sna)
```
---

#Big task:

(apart from conquering the universe)

complete a small project in R using social network analysis, using provided dataset

#Smaller tasks:

- see each topic!
---
#`igraph` package

['igraph'](https://igraph.org/) is a collection for creating and manipulating graphs and analyzing networks. 

- written in C and also exists as Python and R packages.
---
#`statnet` package

['statnet'](http://statnet.org/) is a collection of R packages for 

- management, 

- exploration, 

- statistical analysis, 

- simulation and 

- vizualization of network data. 

The statistical modeling framework is based on [Exponential-family Random Graph Models (ERGMs)](https://en.wikipedia.org/wiki/Exponential_random_graph_models).  
-----

Questions to ask about a network:

- How large the network is?

- How densely connected is it? 

- Is the network made up of one or more distinct groups?

- How compact is the network?

- How clustered are the network members?

---
Size

= the number of members (e.g nodes, vertices or actors)

```{r eval=FALSE}
network.size()

```






---
Resources:

[`statnet` workshop](https://github.com/statnet/Workshops/wiki)

[Awesome list of resources](https://github.com/briatte/awesome-network-analysis)


