---
title: "Practical Applications of Network Science: Workshop "
subtitle: "⚔<br/>with xaringan"
author: "Dr Maria Prokofieva"
institute: "VU Business School"
date: "2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"

      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center")
```
#Practical Applications of Network Science: Workshop:


theory and practical applications of network science to solve today's challenges
---
#Packages to use

[igraph](https://igraph.org/r/)

[statnet](https://cran.r-project.org/web/packages/statnet/)

[sna](https://cran.r-project.org/web/packages/sna/)

```{r}
#load libraries
library (igraph)
library (statnet)
library (sna)
```
---

#Big task:

(apart from conquering the universe)

complete a small project in R using social network analysis, using provided dataset

#Smaller tasks:

- see each topic!
---
#`igraph` package

['igraph'](https://igraph.org/) is a collection for creating and manipulating graphs and analyzing networks. 

- written in C and also exists as Python and R packages.

- [`igraph` package](https://igraph.org/r/)
---
#`statnet` package

['statnet'](http://statnet.org/) is a collection of R packages for 

- management, 

- exploration, 

- statistical analysis, 

- simulation and 

- vizualization of network data. 

- based on [Exponential-family Random Graph Models (ERGMs)](https://en.wikipedia.org/wiki/Exponential_random_graph_models).  
---

Questions to ask about a network:

- How large the network is?

- How densely connected is it? 

- Is the network made up of one or more distinct groups?

- How compact is the network?

- How clustered are the network members?


---

#Dataset

`networkdata` package: a large collection of dataset for network science from wide variety of areas

```{r}
#install.packages("remotes")
#remotes::install_github("schochastics/networkdata")
library(networkdata)
```

```{r eval=FALSE}

#to see the list of datasets
data(package = "networkdata"))
```

```{r}
library(UserNetR)
data(Moreno)

data(flo) #dataset from network package (statnet) 
flo_network<-network(flo, directed=FALSE)

```
---
#Basic concepts

- nodes

**edges**: observed ties in a network (also called edges, arcs, or relations)

- can be directed or undirected

An **undirected tie** is a tie with no direction (e.g. collaborarion)

A **directed tie** is a tie with the direction (e.g. borrowing)

---
#Network characeristics

```{r}
summary(flo_network)
```


---
Size

= the number of members (e.g nodes, vertices or actors)```{r eval=FALSE}
network.size()```{r eval= FALSE}

```{r}
network.size(Moreno)
network.size(flo_network)
```
---
# Density 

proportion of observed ties in a network to the maximum number of possible ties.

a ratio that can range from 0 to 1.

more interconnected networks have density closer to 1.

- to calculate

directed network: 

Number of observed ties
, the maximum number of possible ties among k actors is
k ∗ (k−1), so the formula for density is:
L
k×(k−1)
,
where L is the number of observed ties in the network.

---
---






---
Resources:

[`statnet` workshop](https://github.com/statnet/Workshops/wiki)


Demos are available```{r}
demo(package="igraph")``````{r}
```



[Awesome list of resources](https://github.com/briatte/awesome-network-analysis)
