---
title: "Practical Applications of Network Science: Workshop "
subtitle: "⚔<br/>with xaringan"
author: "Dr Maria Prokofieva"
institute: "VU Business School"
date: "2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"

      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
      
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["shinobi", "ninjutsu", "css/slides.css"]
    seal: true
    self_contained: false
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false      
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center")
```
#Practical Applications of Network Science: Workshop:


Theory and practical applications of network science to solve today's challenges
---
#Packages to use

[igraph](https://igraph.org/r/)

[statnet](https://cran.r-project.org/web/packages/statnet/)

```{r}
#install.packages("igraph","statnet", "ggraph", "tidygraph")

#load libraries
library (igraph)
library (statnet)
library(ggraph)
library(tidygraph)
```
---

# Big task:

(apart from conquering the universe)

complete a small project in R using social network analysis, using the provided dataset

# Smaller tasks:

- see each topic!
---
# Network science:


---
# Why Network Science?

understanding how people, entities, structures are connected

how ... interact

how ... that influence the change in ... across space and time

**Overall**: use network analysis if you need to understand the **system**: 
  
  = how **interactions** or **connections** between **components** make up a **system**


IMAGE
---
# Why Network Science?

If you have any of these questions:

- What are the elements in the system?

- What are the role of those elements in the system

- What are the processes/interactions that build systems?

- What are the dynamics of changes in the system?(e.g. spread of the disease, etc.)
---
# Applications 

social networks - any!

- humans and animals

- ecological communities and populations

- personality traits

- biological systems: genes regulation systems, metabolic systems

- ANY SYSTEM 
---
# Why R?

- free, cross-platform, open-source programming language

- large community of users and large community of academic users

- dynamic growth with new and improved packages

- Comprehensive R Archive Network (CRAN): repository with R packages that have been vetted and deposited on a server
---
# R Packages

This tutorial we will use:

- `igraph`

- `statnet`

Also
- `tidygraph`

- `ggraph` 

Let's talk about them briefly to dig into more details in the practical part.
---
#`igraph` package

['igraph'](https://igraph.org/) is a collection for creating and manipulating graphs and analyzing networks. 

- written in C and also exists as Python and R packages.

- [`igraph` package](https://igraph.org/r/)
---
#`statnet` package

['statnet'](http://statnet.org/) is a collection of R packages for 

- management, 

- exploration, 

- statistical analysis, 

- simulation and 

- vizualization of network data. 

- based on [Exponential-family Random Graph Models (ERGMs)](https://en.wikipedia.org/wiki/Exponential_random_graph_models).  
---
#`ggraph`

[`ggraph`](https://cran.r-project.org/web/packages/ggraph/) is an extension of `ggplot2` for graph visualizations

- use the same layer by layer philosophy

---
# `tidygraph`

[`tidygraph`] (https://cran.r-project.org/web/packages/tidygraph/) is a tidy API for graph manipulation

- Provides a way for dplyr to interact with graphs
---
# Datasets

[`networkdata` package](http://networkrepository.com/): a large collection of dataset for network science from wide variety of areas

```{r}
#install.packages("remotes")
#remotes::install_github("schochastics/networkdata")
library(networkdata)
```

```{r eval=FALSE}

#to see the list of datasets
data(package = "networkdata")
```

---
# What is a network?

Graph theory

A **graph** is made up of **vertices** (also called **nodes**) that are connected by **edges** (also called **links** or relationships).

IMAGE
---
# Edges 

E: observed ties in a network (also called edges, arcs, or relations)

- can be directed or undirected

An **undirected tie** is a tie with no direction (e.g. collaborarion)

A **directed tie** is a tie with the direction (e.g. borrowing) = *arcs*!

graphDir.png

graphUnd.png

---
#Weigthed graph
A directed network where the edges and nodes have different weights, as indicated by their sizes.

IMAGE
graphW.png

--


Single-Mode and Multi-Mode

---
# Questions to ask about a network:

- How large the network is?

- How densely connected is it? 

- Is the network made up of one or more distinct groups?

- How compact is the network?

- How clustered are the network members?

---
# Measures

Graph measures

Node measures

Edge measures

Network measure

---
graphex1.jpg

Network of connections between devices within the Internet. Courtesy of Steve Jurvetson.http://www.flickr.com/photos/jurvetson/916142/
---
#Network characteristics

```{r}
summary(flo_network)
```
---
#Graph measures

**Size**

= the number of members (e.g nodes, vertices or actors)

Some books may refer to the number of nodes, while the number of edges is less common.


```{r eval=FALSE}
network.size()```{r eval= FALSE}

```{r}
network.size(Moreno)
network.size(flo_network)
```
---
# Graph measures

**Density** 

proportion of observed ties in a network to the maximum number of possible ties.

a ratio that can range from 0 to 1.

more interconnected networks have density closer to 1.

---
# Graph measures

**Density** 

Density of directed network:

(directed network is a network with a direction)

##IMAGE DIRECTED NETWORK

Number of observed ties `L` to the maximum number of possible ties among `k` actors `k ∗ (k−1)`, so the formula for density is:

`density= L/(k*k-1))

Loops are excluded! (= a tie between a node and itself)
----
# Graph measures

**Density** 

Density of undirected network

##IMAGE UNDIRECTED NETWORK
,
where L is the number of observed ties in the network.

---
# Graph measures

**Diameter**

The maximum of the shortest distances between all node-pairs in the graph.

---
# Graph measures

## Shortest path 

- unweighted distance: the number of edges between two nodes

-  weighted distance: the sum of weights between two nodes.

IMAGE
---
# Graph measures

## Shortest path algorithms 

**Breadth-First Search(BFS)**: Start at the root node and explore all of the neighbor nodes at the present depth prior to moving on to the nodes at the next level.

**Depth-First Search(DFS)**: Start at the root node and explore as far as possible along each branch before moving on to the branches at the next level.

IMAGE
---
# Node Measure

**Degree Centrality**: Number of edges that a node has 
= the number of links incident upon a node.

**Eigenvector Centrality**: 
---
# Node Measure

**Eigencentrality** is a measure of the importance of a node in a network. 

- assigns relative scores to all nodes in the network that connections to high-scoring nodes contribute more to the score of the node in question than equal connections to low-scoring nodes.

Node high on eigenvector centrality is essentially connected to many high degree nodes.

---
# Node Measure

**Closeness Centrality** 

In a connected graph, the **normalized closeness** of a node is the average length of the shortest path between the node and all other nodes in the graph. 

The more central a node is, the closer it is to all other nodes.

---
# Node Measure

**Betweenness Centrality** 

The **normalized betweenness** of a node is the number of times a node acts as a bridge along the shortest path between two other nodes.

---
# Node Measure

IMAGE

centrality.png
---
# Network Measure

**Subgraphs**: 

Subset of nodes in a network, along with all of the edges linking these nodes.

**Components** 

A subgraph of the network that is disconnected from each other.

---
# Network Measure

**Subgraphs** and **Components** 

IMAGE components.png

---
# Network Measure

**Islands** 

Only retain edges that have weight larger than threshold.

IMAGE islands.png
---
# Network Measure

**Ego Network**

Subnetworks that are centered on a certain node (the ego node) by running a breadth-first search starting at the ego node, and limiting depth smaller than 3.

IMAGE islands.png
---
# Network Measure

**Clustering coefficient**

Measure the proportion of the ego’s friends that are also friends with each other and the degree to which nodes in a graph tend to cluster together.

---
# Network Measure

**Cliques**

Define as a maximal completely connected subgraph of an undirected graph that every two distinct nodes in the clique are adjacent.

---
# Adjacency matrix 

unweighted networks of N nodes without multiple connections = N×N Adjacency matrix

IMAGE
---
# How to store information about the network?

- rectangular data structures?

= rows are observations, columns are variables?

---

IMAGE WITH NETWORK EXAMPLE AND QUESTION TO STORE IT


PROBLEM: how to fit in relational structure?
---
#Network data structure

observation/variable -> tie (=relationship between members of a network)

---
# Sociomatrices

- a matrix containing network information

- a square matrix where a 1 indicates a tie between two nodes, and a 0 indicates no tie

- also called **adjacency matrix**:  1s in the cells = nodes are adjacent to one another in the network

IMAGE
----
# Sociomatrices

Non-directed network: the sociomatrix is symmetric around the diagonal. 

Directed network: EXAMPLE

Simple networks(no self-ties):diagonals are all zeros
---
# Sociomatrices

EXAMPLE

Disadvantages: sparse sociomatrices with large networks
---
# Edge list

=lists every tie in the network. 

= Each row = a single tie from the node listed in the first column to the node listed in the second column.

- much more efficient for large networks

IMAGE
---

# Network Data objects

Types of information:

**Nodes**: list of nodes in network + node labels (Required)

in `statnet` =vertices

**Ties**: list of ties in the network (Required)

**Node attributes**: attributes of the nodes (Optional)


**Tie attributes**: attributes of the ties (Optional)

**Metadata**: other information about the entire network
`statnet` = global information about the network (e.g. directed or not directed, are loops allowed, etc.)
---
# Things to do with networks

Create or import networks



---
# Network in R

- `statnet`: collection of packages to work with networks [`statnet`](http://statnet.org/)

[CRAN](https://cran.r-project.org/package=statnet)

- `igraph`:  collection of network analysis tools with the emphasis on efficiency, portability and ease of use - [`igraph`](https://igraph.org/)

[CRAN](https://cran.r-project.org/web/packages/igraph/index.html)
---
#`igraph` package
```{r}
#install.packages("igraph") #install igraph package if you are yet to install it
library(igraph) #load igraph package
```

---
#Creating network objects in `igraph``

Let's create an *undirected* graph with 3 edges: 1–>2, 2–>3, 3–>1.

```{r}
g1 <- graph( edges=c(1,2, 2,3, 3, 1), n=3, directed=F)
class(g1)
g1
plot(g1) # simple plot of the network
```
---
Let's do more!

```{r}
g2 <- graph( edges=c(1,2, 2,3, 3, 1), n=10, directed=F)
class(g2)
g2
plot(g2) # how does it look like?

```
----
# `make_graph` = `graph`

```{r eval=FALSE}
make_graph(
  edges, #a vector defining the edges, OR character scallar(notable graph - case insensitive!)
  ...,
  n = max(edges),
  isolates = NULL, #for symbolic edge lists
  directed = TRUE #directed or not
)
```

---
#Notable Graphs in `igraph`: 

- `Bull`

The bull graph, 5 vertices, 5 edges, resembles to the head of a bull if drawn properly.

- `Chvatal`

The smallest triangle-free graph that is both 4-chromatic and 4-regular. 
According to the Grunbaum conjecture there exists an m-regular, m-chromatic graph with n vertices for every m>1 and n>2. The Chvatal graph is an example for m=4 and n=12. It has 24 edges.
---
#Notable Graphs in `igraph`: 

- `Coxeter`

A non-Hamiltonian cubic symmetric graph with 28 vertices and 42 edges.

` `Cubical`

The Platonic graph of the cube. A convex regular polyhedron with 8 vertices and 12 edges.

- `Diamond`

A graph with 4 vertices and 5 edges, resembles to a schematic diamond if drawn properly.
---
#Notable Graphs in `igraph`: 

- `Dodecahedral`, `Dodecahedron`

Platonic solid with 20 vertices and 30 edges.

- `Folkman`

The semisymmetric graph with minimum number of vertices, 20 and 40 edges. A semisymmetric graph is regular, edge transitive and not vertex transitive.

---
#Notable Graphs in `igraph`: 

- `Franklin`

This is a graph whose embedding to the Klein bottle can be colored with six colors, it is a counterexample to the neccessity of the Heawood conjecture on a Klein bottle. It has 12 vertices and 18 edges.

- `Frucht`

The Frucht Graph is the smallest cubical graph whose automorphism group consists only of the identity element. It has 12 vertices and 18 edges.
---
#Notable Graphs in `igraph`: 

- `Grotzsch`

The Groetzsch graph is a triangle-free graph with 11 vertices, 20 edges, and chromatic number 4. It is named after German mathematician Herbert Groetzsch, and its existence demonstrates that the assumption of planarity is necessary in Groetzsch's theorem that every triangle-free planar graph is 3-colorable.
---
#Notable Graphs in `igraph`: 

- `Heawood`

The Heawood graph is an undirected graph with 14 vertices and 21 edges. The graph is cubic, and all cycles in the graph have six or more edges. Every smaller cubic graph has shorter cycles, so this graph is the 6-cage, the smallest cubic graph of girth 6.
---
#Notable Graphs in `igraph`: 

- `Herschel`
The Herschel graph is the smallest nonhamiltonian polyhedral graph. It is the unique such graph on 11 nodes, and has 18 edges.
---
#Notable Graphs in `igraph`: 

-` House`
The house graph is a 5-vertex, 6-edge graph, the schematic draw of a house if drawn properly, basicly a triangle of the top of a square.

---
#Notable Graphs in `igraph`: 

- `HouseX`
The same as the house graph with an X in the square. 5 vertices and 8 edges.

---
#Notable Graphs in `igraph`: 

- `Icosahedral`, `Icosahedron`

A Platonic solid with 12 vertices and 30 edges.

- `Krackhardt kite`

A social network with 10 vertices and 18 edges. Krackhardt, D. Assessing the Political Landscape: Structure, Cognition, and Power in Organizations. Admin. Sci. Quart. 35, 342-369, 1990.
---
#Notable Graphs in `igraph`: 

- `Levi`

The graph is a 4-arc transitive cubic graph, it has 30 vertices and 45 edges.

- `McGee`

The McGee graph is the unique 3-regular 7-cage graph, it has 24 vertices and 36 edges.
---
#Notable Graphs in `igraph`: 

- `Meredith`

The Meredith graph is a quartic graph on 70 nodes and 140 edges that is a counterexample to the conjecture that every 4-regular 4-connected graph is Hamiltonian.
---
#Notable Graphs in `igraph`: 

- `Noperfectmatching`

A connected graph with 16 vertices and 27 edges containing no perfect matching. A matching in a graph is a set of pairwise non-adjacent edges; that is, no two edges share a common vertex. A perfect matching is a matching which covers all vertices of the graph.

---
#Notable Graphs in `igraph`: 

- `Nonline`

A graph whose connected components are the 9 graphs whose presence as a vertex-induced subgraph in a graph makes a nonline graph. It has 50 vertices and 72 edges.
---
#Notable Graphs in `igraph`: 

- `Octahedral`, `Octahedron`

Platonic solid with 6 vertices and 12 edges.

- `Petersen`

A 3-regular graph with 10 vertices and 15 edges. It is the smallest hypohamiltonian graph, ie. it is non-hamiltonian but removing any single vertex from it makes it Hamiltonian.
---
#Notable Graphs in `igraph`: 

- `Robertson`

The unique (4,5)-cage graph, ie. a 4-regular graph of girth 5. It has 19 vertices and 38 edges.

- `Smallestcyclicgroup`

A smallest nontrivial graph whose automorphism group is cyclic. It has 9 vertices and 15 edges.
---
#Notable Graphs in `igraph`: 

- `Tetrahedral`, `Tetrahedron`

Platonic solid with 4 vertices and 6 edges.
---
#Notable Graphs in `igraph`: 

- `Thomassen`

The smallest hypotraceable graph, on 34 vertices and 52 edges. A hypotracable graph does not contain a Hamiltonian path but after removing any single vertex from it the remainder always contains a Hamiltonian path. A graph containing a Hamiltonian path is called tracable.
---
#Notable Graphs in `igraph`: 

- `Tutte`

Tait's Hamiltonian graph conjecture states that every 3-connected 3-regular planar graph is Hamiltonian. This graph is a counterexample. It has 46 vertices and 69 edges.

- `Uniquely3colorable`
Returns a 12-vertex, triangle-free graph with chromatic number 3 that is uniquely 3-colorable.
---
#Notable Graphs in `igraph`: 

- `Walther`
An identity graph with 25 vertices and 31 edges. An identity graph has a single graph automorphism, the trivial one.
---
#Notable Graphs in `igraph`: 

- `Zachary`
Social network of friendships between 34 members of a karate club at a US university in the 1970s. See W. W. Zachary, An information flow model for conflict and fission in small groups, Journal of Anthropological Research 33, 452-473 (1977).

---

```{r}
solids <- list(make_graph("Tetrahedron"),
               make_graph("Cubical"),
               make_graph("Octahedron"),
               make_graph("Dodecahedron"),
               make_graph("Icosahedron"))
plot(solids[[2]]) #Cubical graph
```
---
# Graph structure information

`vcount` and `ecount` return the number of vertex. 

`neighbors` returns the number of neighbors for the vertex. is.directed and are.connected return boolean constants. You can use these function to know the type of graph (directed or undirected) and to determine whether there is a link between vertices v1and v2 in graph .get.edgelist function returns the list of edges in a graph. To obtain the adjacency matrix of a graph use get.adjacency function (you also need to install package ?Matrix?in this case).

---
10 minutes + 10 minutes exercise
statnet -creating network object and managing node and tie attributes

10 minutes + 10 minutes exercise
igraph - same

2.2. importing network data and common network data tasks

10 minutes + 10 minutes exercise
filtering network (based on vertex or edge attribute values)

break - 10 minutes

---
##


---
#Classification of graphs

deterministic and stochastic

---
# Creating graphs


---
# Creating graphs - `igraph`

Many functions for creating graphs:

stochastic graph constructors are called ‘games’.
---

#Functions for creating graphs

`graph_from_literal`: create small graphs with a given structure
- R's formula interface, its manual page contains many examples.

- `graph`:  takes numeric vertex ids directly

- `graph.atlas`: create graph from the Graph Atlas

- `make_graph`: create some special graphs.

---
#Functions for creating graphs from field data
To create graphs from field data, graph_from_edgelist, graph_from_data_frame and graph_from_adjacency_matrix are probably the best choices.

The igraph package includes some classic random graphs like the Erdos-Renyi GNP and GNM graphs (sample_gnp, sample_gnm) and some recent popular models, like preferential attachment (sample_pa) and the small-world model (sample_smallworld).





---
Resources:

[`statnet` workshop](https://github.com/statnet/Workshops/wiki)


Demos are available```{r}
demo(package="igraph")``````{r}
```



[Awesome list of resources](https://github.com/briatte/awesome-network-analysis)
