---
title: 'Case Study: Twitter'
author: "Maria Prokofieva"
date: "19/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using network analysis for Twitter data

Social media networks is the most "networks" that come to mind when I mentioned network analysis to "ordinary" people.

So let's dig into them!

In this case study we will focus on 

**Network Models**: develop a model that describes the relationships between users

**Key Players**: identify the most influential users in the network based on a different context

**Tie Strength**: measure how strong the relationships are 

**Network Cohesion**: To measure how cohesive entities in the networks towards network behavior.


We are going to use our main libraries, so let's load them 

```{r}
#install.packages("tidyverse")
#install.packages("igraph")



library(igraph)
library(ggraph)
library(tidygraph)
```

```{r}
# for data wrangling. very helpfull for preparing nodes and edges data
library(tidyverse) 
library(lubridate)

# for building network and visualization 
library(tidygraph)
library(graphlayouts)
# already included in tidygraph but just fyi

#these packages are already included in tidygraph, but just for reference

library(igraph)
library(ggraph)

# for crawling Twitter data 
library(rtweet)
# for visualizing
library(extrafont)
loadfonts(device = "pdf") 
```

To get the network model we will be creating a graph of the Twitter users for the collected data

To get the Key Players in the network we will use the meausures of **centrality**.


Now, let's startw with loading the twitter data. 

We can use [`rtweet` package](https://cran.r-project.org/web/packages/rtweet/) to download it (and we are going to do this later!) but at this stage - let's work with the collected `.csv` file

```{r}
tweets <- read.csv("http://nodeassets.nbcnews.com/russian-twitter-trolls/tweets.csv",
                   stringsAsFactors = F, sep = ",")
```


```{r}
# gather teamalgoritma data
user_Twitter <- lookup_users("rstudio")
# get RStudio followers and its account details
followers <- get_followers("rstudio",n = user_Twitter$followers_count,retryonratelimit = T)
detail_followers <- lookup_users(followers$user_id)
detail_followers <- data.frame(lapply(detail_followers,as.character),
stringsAsFactors = F)
```

This data is on EXTEND 


```{r}
active_fol <- detail_followers %>% select(user_id,screen_name,created_at,followers_count,friends_count,favourites_count) %>%
  mutate(created_at = ymd_hms(created_at),
         followers_count = as.numeric(followers_count),
         friends_count = as.numeric(friends_count),
         favourites_count = as.numeric(favourites_count)) %>%
  filter((followers_count > 100 & followers_count < 6000), friends_count > 75, favourites_count > 10, 
         created_at > "2020-03-15") %>%
  arrange(-followers_count)
```

```{r}
flt_n <- function(x){
  if(x > 1500){
    x*0.5
  }else{x*0.75}
}
```


```{r}
# Create empty list and name it after their screen name
follower <- vector(mode = 'list', length = length(active_fol$screen_name))
names(follower) <- active_fol$screen_name
# 
for (i in seq_along(active_fol$screen_name)) {
  message("Getting followers for user #", i, "/130")
  follower[[i]] <- get_followers(active_fol$screen_name[i], 
                                  n = round(flt_n(active_fol$followers_count[i])), 
                                retryonratelimit = TRUE)
  
  if(i %% 5 == 0){
    message("sleep for 5 minutes")
    Sys.sleep(5*60)
    } 
}
```


```{r}
# convert list to dataframe
follower_x <- bind_rows(foler, .id = "screen_name")
active_fol_x <- active_fol %>% select(user_id,screen_name)
# left join to convert screen_name into its user id
follower_join <- left_join(follower_x, active_fol_x, by="screen_name")
# subset to new dataframe with new column name and delete NA
follower <- follower_join %>% select(user_id.x,user_id.y) %>%
  setNames(c("follower","active_user")) %>% 
  na.omit()
```

