---
title: 'Practical Applications of Network Science: Workshop'
author: "Maria Prokofieva"
date: "06/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(igraph)
library(statnet)
library(tidygraph)
library(ggraph)
```

# Practical Applications of Network Science: Workshop

This two-day workshop will include:

2 days * 3hrs hands-on workshops in R

Target: real-world applications of network science

**Skill level**: introductory level 

- No knowledge/basic knowledge of network science

- Basic knowledge of R and `tidyverse`

**Packages used**

[`igraph`](https://igraph.org/r/)

[`statnet`](https://cran.r-project.org/web/packages/statnet/)

[`ggraph`](https://cran.r-project.org/web/packages/ggraph/)

[`tidygraph`](https://cran.r-project.org/web/packages/tidygraph/) 

*Datasets*


[TidyTuesday French Train dataset]
(https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-26/small_trains.csv")


### Resources

[`statnet` workshop](https://github.com/statnet/Workshops/wiki)

[Awesome list of resources](https://github.com/briatte/awesome-network-analysis)

## Tasks

1. Install (if not yet installed) and load the required packages

```{r}
#install.packages("igraph", "statnet", "ggraph", "tidygraph")
library(tidyverse)
library(igraph, statnet)
```

## 2. `Igraph`
###Creating network objects from files

####Dataset

EXPAND

```{r}
small_trains <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-26/small_trains.csv")

head(small_trains$departure_station)

```
####Data preparation

Let's create the **node list**:

We need to isolate `distinct` stations from both the “departure_station” and “arrival_station” columns and then join the information from these columns together.

```{r}
routes_from <- small_trains %>%
  distinct(departure_station) %>%
  rename(label = departure_station)

departure_station

routes_to <- small_trains %>%
  distinct(arrival_station) %>%
  rename(label = arrival_station)
```

And use a full join to include all unique places from both the departure_station and arrival_station.

```{r}
nodes <- full_join(routes_from, routes_to, by = "label")
head(nodes)
```
Let's give unique IDs for each station

```{r}
nodes <- nodes %>% rowid_to_column("id")
head(nodes)
```

Let's create the **edge list**:

```{r}
per_route <- small_trains %>%  
  group_by(departure_station, arrival_station) %>%
  summarise(weight = mean(total_num_trips)) %>% 
  ungroup()
head(per_route)
```
But... we still have “departure_station” and “arrival_station” columns contain labels rather than IDs. 

Let's link the IDs that have been assigned in nodes to each station in both the “departure_station” and “arrival_station” columns.

```{r}
edges <- per_route %>% 
  left_join(nodes, by = c("departure_station" = "label")) %>% 
  rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("arrival_station" = "label")) %>% 
  rename(to = id)

head(edges)
```
Let's reorder the columns so that the dataset starts with “from” and “to” (IDs). We also do not need “departure_station” and “arrival_station” columns with the names of the stations.

Let's use `select() function` to include the “from”, “to”, and “weight” columns.

```{r}
edges <- edges%>%
  select(from, to, weight)

head(edges)
```
Now `edges` and `nodes` provides us with all of the information necessary to create network objects with the `igraph` and `statnet`.


###`statnet` - `network` object

To create a network object in `statnet` we use `network` function which comes with the following options:

```{r eval=FALSE}
network(x, vertex.attr = NULL, vertex.attrnames = NULL,
  directed = TRUE, hyper = FALSE, loops = FALSE, multiple = FALSE,
  bipartite = FALSE, ...)
```

where

- `x`: a matrix giving the network structure in adjacency, incidence, or edgelist form

- `vertex.attr`:	a list containing vertex attributes

- `directed`: `TRUE` by default - should edges be interpreted as directed?

- `loops`: `FALSE` by default - are hyperedges allowed?

- `matrix.type`: 	
one of "adjacency", "edgelist", "incidence"
?
```{r}
routes_network <- network(edges, vertex.attr = nodes, matrix.type = "edgelist", ignore.eval = FALSE)

```

Let's see the class of the object:
```{r}
class(routes_network)

```

#### `igraph` object

Let's convert the data to an `igraph` object. The format for the is:



```{r eval=FALSE}
net <- graph_from_data_frame(d=links, vertices=nodes, directed=T) 
```

The `graph.data.frame` function, which takes two data frames: `d` and `vertices`.

- `d` describes the edges of the network - it should start with two columns containing the source and target node IDs for each network tie.

- `vertices` should start with a column of node IDs.

- Any additional columns in the data frame are interpreted as attributes.

# Examine the resulting object:
class(net)
net 

# We can look at the nodes, edges, and their attributes:
E(net)
V(net)
E(net)$type
V(net)$media

plot(net, edge.arrow.size=.4,vertex.label=NA)

# Removing loops from the graph:
net <- simplify(net, remove.multiple = F, remove.loops = T) 

# If you need them, you can extract an edge list or a matrix from igraph networks.
as_edgelist(net, names=T)
as_adjacency_matrix(net, attr="weight")

# Or data frames describing nodes and edges:
as_data_frame(net, what="edges")
as_data_frame(net, what="vertices")


#  ------->> DATASET 2 --------

head(nodes2)
head(links2)

net2 <- graph_from_incidence_matrix(links2)

# A built-in vertex attribute 'type' shows which mode vertices belong to.
table(V(net2)$type)

plot(net2,vertex.label=NA)

# To transform a one-mode network matrix into an igraph object,
# use graph_from_adjacency_matrix()

# We can also easily generate bipartite projections for the two-mode network:
# (co-memberships are easy to calculate by multiplying the network matrix by
# its transposed matrix, or using igraph's bipartite.projection function)

net2.bp <- bipartite.projection(net2)

# We can calculate the projections manually as well:
#   as_incidence_matrix(net2)  %*% t(as_incidence_matrix(net2))
# t(as_incidence_matrix(net2)) %*%   as_incidence_matrix(net2)

plot(net2.bp$proj1, vertex.label.color="black", vertex.label.dist=1,
     vertex.label=nodes2$media[!is.na(nodes2$media.type)])

plot(net2.bp$proj2, vertex.label.color="black", vertex.label.dist=1,
     vertex.label=nodes2$media[ is.na(nodes2$media.type)])


# ================ 5. Plotting networks with igraph ================

 
#  ------->> Plot parameters in igraph --------

# Plotting with igraph: node options (starting with 'vertex.') and edge options
# (starting with 'edge.'). A list of options is included in your handout.
?igraph.plotting

# We can set the node & edge options in two ways - one is to specify
# them in the plot() function, as we are doing below.

# Plot with curved edges (edge.curved=.1) and reduce arrow size:
plot(net, edge.arrow.size=.4, edge.curved=.1)

# Set node color to orange and the border color to hex #555555
# Replace the vertex label with the node names stored in "media"
plot(net, edge.arrow.size=.2, edge.curved=0,
     vertex.color="orange", vertex.frame.color="#555555",
     vertex.label=V(net)$media, vertex.label.color="black",
     vertex.label.cex=.7) 

# The second way to set attributes is to add them to the igraph object.

# Generate colors based on media type:
colrs <- c("gray50", "tomato", "gold")
V(net)$color <- colrs[V(net)$media.type]

# Set node size based on audience size:
V(net)$size <- V(net)$audience.size*0.7

# The labels are currently node IDs.
# Setting them to NA will render no labels:
V(net)$label.color <- "black"
V(net)$label <- NA

# Set edge width based on weight:
E(net)$width <- E(net)$weight/6

#change arrow size and edge color:
E(net)$arrow.size <- .2
E(net)$edge.color <- "gray80"

plot(net) 

# We can also override the attributes explicitly in the plot:
plot(net, edge.color="orange", vertex.color="gray50") 


# We can also add a legend explaining the meaning of the colors we used:
plot(net) 
legend(x=-1.1, y=-1.1, c("Newspaper","Television", "Online News"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2.5, bty="n", ncol=1)


# Sometimes, especially with semantic networks, we may be interested in 
# plotting only the labels of the nodes:

plot(net, vertex.shape="none", vertex.label=V(net)$media, 
     vertex.label.font=2, vertex.label.color="gray40",
     vertex.label.cex=.7, edge.color="gray85")


# Let's color the edges of the graph based on their source node color.
# We'll get the starting node for each edge with "ends()".
edge.start <- ends(net, es=E(net), names=F)[,1]
edge.col <- V(net)$color[edge.start]

plot(net, edge.color=edge.col, edge.curved=.1)


#  ------->> Network Layouts --------

# Network layouts are algorithms that return coordinates for each
# node in a network.

# Let's generate a slightly larger 80-node graph.

net.bg <- sample_pa(80, 1.2) 
V(net.bg)$size <- 8
V(net.bg)$frame.color <- "white"
V(net.bg)$color <- "orange"
V(net.bg)$label <- "" 
E(net.bg)$arrow.mode <- 0
plot(net.bg)

# You can set the layout in the plot function:
plot(net.bg, layout=layout_randomly)

# Or calculate the vertex coordinates in advance:
l <- layout_in_circle(net.bg)
plot(net.bg, layout=l)

# l is simply a matrix of x,y coordinates (N x 2) for the N nodes in the graph. 
# You can generate your own:
l
l <- cbind(1:vcount(net.bg), c(1, vcount(net.bg):2))
plot(net.bg, layout=l)

# This layout is just an example and not very helpful - thankfully
# igraph has a number of built-in layouts, including:

# Randomly placed vertices
l <- layout_randomly(net.bg)
plot(net.bg, layout=l)

# Circle layout
l <- layout_in_circle(net.bg)
plot(net.bg, layout=l)

# 3D sphere layout
l <- layout_on_sphere(net.bg)
plot(net.bg, layout=l)

# The Fruchterman-Reingold force-directed algorithm 
# Nice but slow, most often used in graphs smaller than ~1000 vertices. 
l <- layout_with_fr(net.bg)
plot(net.bg, layout=l)

# You will also notice that the layout is not deterministic - different runs 
# will result in slightly different configurations. Saving the layout in l
# allows us to get the exact same result multiple times.
par(mfrow=c(2,2), mar=c(1,1,1,1))
plot(net.bg, layout=layout_with_fr)
plot(net.bg, layout=layout_with_fr)
plot(net.bg, layout=l)
plot(net.bg, layout=l)
dev.off()

# By default, the coordinates of the plots are rescaled to the [-1,1] interval
# for both x and y. You can change that with the parameter "rescale=FALSE"
# and rescale your plot manually by multiplying the coordinates by a scalar.
# You can use norm_coords to normalize the plot with the boundaries you want.

# Get the layout coordinates:
l <- layout_with_fr(net.bg)
# Normalize them so that they are in the -1, 1 interval:
l <- norm_coords(l, ymin=-1, ymax=1, xmin=-1, xmax=1)

par(mfrow=c(2,2), mar=c(0,0,0,0))
plot(net.bg, rescale=F, layout=l*0.4)
plot(net.bg, rescale=F, layout=l*0.8)
plot(net.bg, rescale=F, layout=l*1.2)
plot(net.bg, rescale=F, layout=l*1.6)
dev.off()

# Another popular force-directed algorithm that produces nice results for
# connected graphs is Kamada Kawai. Like Fruchterman Reingold, it attempts to 
# minimize the energy in a spring system.

l <- layout_with_kk(net.bg)
plot(net.bg, layout=l)

# The LGL algorithm is for large connected graphs. Here you can specify a root - 
# the node that will be placed in the middle of the layout.
plot(net.bg, layout=layout_with_lgl)

# By default, igraph uses a layout called layout_nicely which selects
# an appropriate layout algorithm based on the properties of the graph. 

# Check out all available layouts in igraph:
?igraph::layout_

layouts <- grep("^layout_", ls("package:igraph"), value=TRUE)[-1] 
# Remove layouts that do not apply to our graph.
layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree", layouts)]

par(mfrow=c(3,3), mar=c(1,1,1,1))

for (layout in layouts) {
  print(layout)
  l <- do.call(layout, list(net)) 
  plot(net, edge.arrow.mode=0, layout=l, main=layout) }

dev.off()


-----------------------------------
# * TASK FOR WORKSHOP PARTICIPANTS:

# Plot the Zachary karate club network with four different layouts of your choice.

-----------------------------------

  

# ------->> Improving network plots --------

plot(net)

# Notice that this network plot is still not too helpful.
# We can identify the type and size of nodes, but cannot see
# much about the structure since the links we're examining are so dense.
# One way to approach this is to see if we can sparsify the network.

hist(links$weight)
mean(links$weight)
sd(links$weight)

# There are more sophisticated ways to extract the key edges,
# but for the purposes of this excercise we'll only keep ones
# that have weight higher than the mean for the network.

# We can delete edges using delete_edges(net, edges)
cut.off <- mean(links$weight) 
net.sp <- delete_edges(net, E(net)[weight<cut.off])
plot(net.sp) 

# Another way to think about this is to plot the two tie types 
# (hyperlik & mention) separately:

E(net)$width <- 2
plot(net, edge.color=c("dark red", "slategrey")[(E(net)$type=="hyperlink")+1],
      vertex.color="gray40", layout=layout_in_circle)

# Another way to delete edges:  
net.m <- net - E(net)[E(net)$type=="hyperlink"]
net.h <- net - E(net)[E(net)$type=="mention"]

# Plot the two links separately:
par(mfrow=c(1,2))

plot(net.h, vertex.color="orange", main="Tie: Hyperlink")
plot(net.m, vertex.color="lightsteelblue2", main="Tie: Mention")

dev.off()

# Make sure the nodes stay in place in both plots:
par(mfrow=c(1,2),mar=c(1,1,4,1))

l <- layout_with_fr(net)
plot(net.h, vertex.color="orange", layout=l, main="Tie: Hyperlink")
plot(net.m, vertex.color="lightsteelblue2", layout=l, main="Tie: Mention")

dev.off()


# ------->> Interactive plotting with tkplot -------- 

# R and igraph offer interactive plotting capabilities
# (mostly helpful for small networks)

tkid <- tkplot(net) #tkid is the id of the tkplot
l <- tkplot.getcoords(tkid) # grab the coordinates from tkplot
tk_close(tkid, window.close = T)
plot(net, layout=l)


# ------->> Heatmaps as a way to represent networks -------- 

# A quick reminder that there are other ways to represent a network:

# Heatmap of the network matrix:
netm <- get.adjacency(net, attr="weight", sparse=F)
colnames(netm) <- V(net)$media
rownames(netm) <- V(net)$media

palf <- colorRampPalette(c("gold", "dark orange")) 
heatmap(netm[,17:1], Rowv = NA, Colv = NA, col = palf(20), 
        scale="none", margins=c(10,10) )


# ------->> Plotting two-mode networks with igraph --------  

head(nodes2)
head(links2)

net2
plot(net2)

# This time we will make nodes look different based on their type.
V(net2)$color <- c("steel blue", "orange")[V(net2)$type+1]
V(net2)$shape <- c("square", "circle")[V(net2)$type+1]
V(net2)$label <- ""
V(net2)$label[V(net2)$type==F] <- nodes2$media[V(net2)$type==F] 
V(net2)$label.cex=.6
V(net2)$label.font=2

plot(net2, vertex.label.color="white", vertex.size=(2-V(net2)$type)*8) 

plot(net2, vertex.label=NA, vertex.size=7, layout=layout_as_bipartite) 

 
# Using text as nodes:
par(mar=c(0,0,0,0))
plot(net2, vertex.shape="none", vertex.label=nodes2$media,
     vertex.label.color=V(net2)$color, vertex.label.font=2, 
     vertex.label.cex=.95, edge.color="gray70",  edge.width=2)

dev.off()


# ================ 6. Network and node descriptives ================


# Density
# The proportion of present edges from all possible ties.
edge_density(net, loops=F)
ecount(net)/(vcount(net)*(vcount(net)-1)) #for a directed network

# Reciprocity
# The proportion of reciprocated ties (for a directed network).
reciprocity(net)
dyad_census(net) # Mutual, asymmetric, and null node pairs
2*dyad_census(net)$mut/ecount(net) # Calculating reciprocity

# Transitivity
# global - ratio of triangles (direction disregarded) to connected triples
# local - ratio of triangles to connected triples each vertex is part of
transitivity(net, type="global")  # net is treated as an undirected network
transitivity(as.undirected(net, mode="collapse")) # same as above
transitivity(net, type="local")
triad_census(net) # for directed networks

# Triad types (per Davis & Leinhardt):
# 
# 003  A, B, C, empty triad.
# 012  A->B, C 
# 102  A<->B, C  
# 021D A<-B->C 
# 021U A->B<-C 
# 021C A->B->C
# 111D A<->B<-C
# 111U A<->B->C
# 030T A->B<-C, A->C
# 030C A<-B<-C, A->C.
# 201  A<->B<->C.
# 120D A<-B->C, A<->C.
# 120U A->B<-C, A<->C.
# 120C A->B->C, A<->C.
# 210  A->B<->C, A<->C.
# 300  A<->B<->C, A<->C, completely connected.


# Diameter (longest geodesic distance)
# Note that edge weights are used by default, unless set to NA.
diameter(net, directed=F, weights=NA)
diameter(net, directed=F)
diam <- get_diameter(net, directed=T)
diam

# Note: vertex sequences asked to behave as a vector produce numeric index of nodes
class(diam)
as.vector(diam)

# Color nodes along the diameter:
vcol <- rep("gray40", vcount(net))
vcol[diam] <- "gold"
ecol <- rep("gray80", ecount(net))
ecol[E(net, path=diam)] <- "orange" 
# E(net, path=diam) finds edges along a path, here 'diam'
plot(net, vertex.color=vcol, edge.color=ecol, edge.arrow.mode=0)

# Node degrees
# 'degree' has a mode of 'in' for in-degree, 'out' for out-degree,
# and 'all' or 'total' for total degree. 
deg <- degree(net, mode="all")
plot(net, vertex.size=deg*3)
hist(deg, breaks=1:vcount(net)-1, main="Histogram of node degree")

# Degree distribution
deg.dist <- degree_distribution(net, cumulative=T, mode="all")
plot( x=0:max(deg), y=1-deg.dist, pch=19, cex=1.2, col="orange", 
      xlab="Degree", ylab="Cumulative Frequency")


# Centrality & centralization

# Centrality functions (vertex level) and centralization functions (graph level).
# The centralization functions return "res" - vertex centrality, "centralization", 
# and "theoretical_max" - maximum centralization score for a graph of that size.
# The centrality functions can run on a subset of nodes (set with the "vids" parameter)

# Degree (number of ties)
degree(net, mode="in")
centr_degree(net, mode="in", normalized=T)

# Closeness (centrality based on distance to others in the graph)
# Inverse of the node's average geodesic distance to others in the network
closeness(net, mode="all", weights=NA) 
centr_clo(net, mode="all", normalized=T) 

# Eigenvector (centrality proportional to the sum of connection centralities)
# Values of the first eigenvector of the graph adjacency matrix
eigen_centrality(net, directed=T, weights=NA)
centr_eigen(net, directed=T, normalized=T) 

# Betweenness (centrality based on a broker position connecting others)
# (Number of geodesics that pass through the node or the edge)
betweenness(net, directed=T, weights=NA)
edge_betweenness(net, directed=T, weights=NA)
centr_betw(net, directed=T, normalized=T)



-----------------------------------
# * TASK FOR WORKSHOP PARTICIPANTS:

# Compute the degree, closeness, eigenvector, and betweenness centrality of
# the actors in the Zachary karate club network. Plot the network, sizing the
# nodes based on the different centrality types.

-----------------------------------

  

# Hubs and authorities

# The hubs and authorities algorithm developed by Jon Kleinberg was initially used 
# to examine web pages. Hubs were expected to contain catalogues with a large number 
# of outgoing links; while authorities would get many incoming links from hubs, 
# presumably because of their high-quality relevant information. 

hs <- hub_score(net, weights=NA)$vector
as <- authority_score(net, weights=NA)$vector

par(mfrow=c(1,2))
 plot(net, vertex.size=hs*50, main="Hubs")
 plot(net, vertex.size=as*30, main="Authorities")
dev.off()


# ================ 7. Distances and paths ================


# Average path length 
# The mean of the shortest distance between each pair of nodes in the network 
# (in both directions for directed graphs). 
mean_distance(net, directed=F)
mean_distance(net, directed=T)

# We can also find the length of all shortest paths in the graph:
distances(net) # with edge weights
distances(net, weights=NA) # ignore weights

# We can extract the distances to a node or set of nodes we are interested in.
# Here we will get the distance of every media from the New York Times.
dist.from.NYT <- distances(net, v=V(net)[media=="NY Times"], to=V(net), weights=NA)

# Set colors to plot the distances:
oranges <- colorRampPalette(c("dark red", "gold"))
col <- oranges(max(dist.from.NYT)+1)
col <- col[dist.from.NYT+1]

plot(net, vertex.color=col, vertex.label=dist.from.NYT, edge.arrow.size=.6, 
     vertex.label.color="white")

# We can also find the shortest path between specific nodes.
# Say here between MSNBC and the New York Post:
news.path <- shortest_paths(net, 
                            from = V(net)[media=="MSNBC"], 
                             to  = V(net)[media=="New York Post"],
                             output = "both") # both path nodes and edges

# Generate edge color variable to plot the path:
ecol <- rep("gray80", ecount(net))
ecol[unlist(news.path$epath)] <- "orange"
# Generate edge width variable to plot the path:
ew <- rep(2, ecount(net))
ew[unlist(news.path$epath)] <- 4
# Generate node color variable to plot the path:
vcol <- rep("gray40", vcount(net))
vcol[unlist(news.path$vpath)] <- "gold"

plot(net, vertex.color=vcol, edge.color=ecol, 
     edge.width=ew, edge.arrow.mode=0)


# Identify the edges going into or out of a vertex, for instance the WSJ.
# For a single node, use 'incident()', for multiple nodes use 'incident_edges()'
inc.edges <- incident(net, V(net)[media=="Wall Street Journal"], mode="all")

# Set colors to plot the selected edges.
ecol <- rep("gray80", ecount(net))
ecol[inc.edges] <- "orange"
vcol <- rep("grey40", vcount(net))
vcol[V(net)$media=="Wall Street Journal"] <- "gold"
plot(net, vertex.color=vcol, edge.color=ecol)


# We can also easily identify the immediate neighbors of a vertex, say WSJ.
# The 'neighbors' function finds all nodes one step out from the focal actor.
# To find the neighbors for multiple nodes, use 'adjacent_vertices()'.
# To find node neighborhoods going more than one step out, use function 'ego()'
# with parameter 'order' set to the number of steps out to go from the focal node(s).

neigh.nodes <- neighbors(net, V(net)[media=="Wall Street Journal"], mode="out")

# Set colors to plot the neighbors:
vcol[neigh.nodes] <- "#ff9d00"
plot(net, vertex.color=vcol)

# Special operators for the indexing of edge sequences: %--%, %->%, %<-%
# E(network)[X %--% Y] selects edges between vertex sets X and Y, ignoring direction
# E(network)[X %->% Y] selects edges from vertex sets X to vertex set Y
# E(network)[X %->% Y] selects edges from vertex sets Y to vertex set X

# For example, select edges from newspapers to online sources:
E(net)[ V(net)[type.label=="Newspaper"] %->% V(net)[type.label=="Online"] ]

# Cocitation (for a couple of nodes, how many shared nominations they have)
cocitation(net)



# ================ 8. Subgroups and communities ================

# Converting 'net' to an undirected network.
# There are several ways to do that: we can create an undirected link between any pair
# of connected nodes (mode="collapse), or create an undirected link for each directed
# one (mode="each"), or create an undirected link for each symmetric link (mode="mutual").
# In cases when A -> B and B -> A are collapsed into a single undirected link, we
# need to specify what to do with the edge attributes. Here we have said that
# the 'weight' of links should be summed, and all other edge attributes ignored.

net.sym <- as.undirected(net, mode="collapse", edge.attr.comb=list(weight="sum", "ignore"))


#  ------->> Cliques --------

 # Find cliques (complete subgraphs of an undirected graph)
cliques(net.sym) # list of cliques       
sapply(cliques(net.sym), length) # clique sizes
largest_cliques(net.sym) # cliques with max number of nodes

vcol <- rep("grey80", vcount(net.sym))
vcol[unlist(largest_cliques(net.sym))] <- "gold"
plot(net.sym, vertex.label=V(net.sym)$name, vertex.color=vcol)



#  ------->> Communities --------

# A number of algorithms aim to detect groups that consist of densely connected nodes
# with fewer connections across groups. 

# Community detection based on edge betweenness (Newman-Girvan)
# High-betweenness edges are removed sequentially (recalculating at each step)
# and the best partitioning of the network is selected.
ceb <- cluster_edge_betweenness(net) 
dendPlot(ceb, mode="hclust")
plot(ceb, net) 

# Let's examine the community detection igraph object:
class(ceb)
length(ceb)     # number of communities
membership(ceb) # community membership for each node
crossing(ceb, net)   # boolean vector: TRUE for edges across communities
modularity(ceb) # how modular the graph partitioning is

# High modularity for a partitioning reflects dense connections within communities 
# and sparse connections across communities.


# Community detection based on propagating labels
# Assigns node labels, randomizes, and replaces each vertex's label with
# the label that appears most frequently among neighbors. Repeated until
# each vertex has the most common label of its neighbors.
clp <- cluster_label_prop(net)
plot(clp, net)

# Community detection based on greedy optimization of modularity
cfg <- cluster_fast_greedy(as.undirected(net))
plot(cfg, as.undirected(net))
 
# We can also plot the communities without relying on their built-in plot:
V(net)$community <- cfg$membership
colrs <- adjustcolor( c("gray50", "tomato", "gold", "yellowgreen"), alpha=.6)
plot(net, vertex.color=colrs[V(net)$community])


-----------------------------------
# * TASK FOR WORKSHOP PARTICIPANTS:

# Plot the results of the three different community detection algorithms 
#  applied to  the Zachary karate club network. 
  
-----------------------------------


# K-core decomposition
# The k-core is the maximal subgraph in which every node has degree of at least k
# This also means that the (k+1)-core will be a subgraph of the k-core.
# The result here gives the coreness of each vertex in the network.
kc <- coreness(net, mode="all")
plot(net, vertex.size=kc*6, vertex.label=kc, vertex.color=colrs[kc])


# ================ 9. Assortativity and Homophily ================

# Assortativity (homophily)
# The tendency of nodes to connect to others who are similar on some variable.
# assortativity_nominal() is for categorical variables (labels)
# assortativity() is for ordinal and above variables
# assortativity_degree() checks assortativity in node degrees

V(net)$type.label
V(net)$media.type

assortativity_nominal(net, V(net)$media.type, directed=F)

assortativity(net, V(net)$audience.size, directed=F)

assortativity_degree(net, directed=F)










--------
---------
---------
-------
-------

Let's look at some of the characteristics of the graph using `igraph`


# Visualization
```{r}
plot(g3, edge.arrow.size=.5, vertex.color="gold", vertex.size=15,
vertex.frame.color="gray", vertex.label.color="black",
vertex.label.cex=0.8, vertex.label.dist=2, edge.curved=0.2)

```



------------

Day 2 - 3 hrs

10 minutes
setup and refresher on the topic - 

10 minutes
2.1.Network Analysis in Social Media Research 

Overview of social media research
SNA applications in social media research and main studies
(differentiation between self-reported and digitally collected network data)
ethical considerations
interpretation of basic metrics

break 10 minutes

2.2. Use case: Twitter analysis - 1 hr
exploratory analysis of the selected dataset, and interpretation of results.

break 10 minutes

1 hr
3.1 Network Analysis in business research - business analytics

Overview of business analytics research - supply chain operations
business analytics research and applications of SNA - amazon dataset

Exploratory analysis of the selected dataset, and interpretation of results.



----
alternative --------------
Introduction to the analysis of socio-technical networks, and applications of network analytic techniques in LA, i.e. community detection, bipartite network analysis, network clustering, integration with text analysis. Presentation of community detection, information flow analysis, and statistical approaches in network analysis. The students will be expected to select one approach out of those presented, and implement it on one of the suggested datasets in R.
---- !!!!!!----------

2.2 Extended feautures of Graphs - 30 minutes

The Page Rank algorithm - Calculating the Google PageRank for the specified vertices.

https://www.r-bloggers.com/2019/05/benchmark-of-popular-graph-network-packages/

[Stanford Large Network Dataset Collection](https://snap.stanford.edu/data/index.html)
[amazon dataset](https://snap.stanford.edu/data/index.html#amazon) 
[google dataset](https://snap.stanford.edu/data/web-Google.html) 

break - 10 minutes

2.3. Other popular packages, benchmarking
30 minutes

graph-tool
igraph
networkit
networkx
snap

2.4. Wrap-up 



