---
title: 'Practical Applications of Network Science: Workshop'
author: "Maria Prokofieva"
date: "06/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)
```

# Practical Applications of Network Science: Workshop

This two-day workshop will include:

2 days * 3hrs hands-on workshops in R

Target: real-world applications of network science


**Skill level**: introductory level 

- No knowledge/basic knowledge of network science

- Basic knowledge of R and `tidyverse`

**Packages used**

[`igraph`](https://igraph.org/r/)

[`statnet`](https://cran.r-project.org/web/packages/statnet/)

[`sna`](https://cran.r-project.org/web/packages/sna/)

*Datasets*

## Tasks

Install (if not yet installed) and load the `igraph` package

```{r}
#install.packages("igraph")
library(igraph)
```

Creating a network object in `igraph`

**Empty graph**:

```{r}
g<-graph.empty(n=10, directed=TRUE)
plot(g)
```
**Complete graph**: each pair of vertices has an edge connecting them, we use `LOOPS = FALSE` to not add self edges

```{r}
g<-graph.full(n=10, directed = FALSE, loops = FALSE)
plot(g)
```

**Stars**
```{r}
g<-graph.star(n=10, mode="out")
plot(g)

g<-graph.star(n=10, mode="in")
plot(g)
```

**Rings**
```{r}
g<-graph.ring(n=10)
```

Now let's create more "complicated" graphs with the given list of edges

```{r}
edges <- c(1,2, 3,2, 2,4)
g<-graph(edges, n=max(edges), directed=TRUE)
plot(g)
```

Can we do a **named graph**? Sure we can!

When you use vertex names in your edge list, you don't need the number of nodes

```{r}
g3 <- graph( c("expresso", "milk", "milk", "chocolate", "cream", "chocolate", "expresso", "cream")) # named vertices 
plot(g3)
```
For smaller graphs we can use

- ` - `for undirected tie, 

- `+-` for pointing left or 

-`-+` for directed ties  & right, 

- `++` for a symmetric tie, and 

- “:” for sets of vertices.

```{r}

plot(graph_from_literal("dogs"---"cats", "cats"---"mice")) # the number of dashes doesn't matter

```
###Working with network attributes, edges and verteces

To access vertices and edges:

```{r}
E(g3) # The edges of the object

V(g3) # The vertices of the object
```
To view the network in a matrix form

```{r}
g3[]
```

Add attributes to the network, vertices, or edges:

```{r}
V(g3)$name # automatically generated when we created the network.

V(g3)$calories <- c("male", "male", "male", "male", "female", "female", "male")
E(g3)$type <- "yumminess" # Edge attribute, assign "yumminess" to all edges
E(g3)$weight <- 10 # Edge weight, setting all existing edges to 10
```
To examine attributes:
```{r}
edge_attr(g3)

```


Let's look at some of the characteristics of the graph using `igraph`


# Visualization
```{r}
plot(g3, edge.arrow.size=.5, vertex.color="gold", vertex.size=15,
vertex.frame.color="gray", vertex.label.color="black",
vertex.label.cex=0.8, vertex.label.dist=2, edge.curved=0.2)

```



------------

Day 2 - 3 hrs

10 minutes
setup and refresher on the topic - 

10 minutes
2.1.Network Analysis in Social Media Research 

Overview of social media research
SNA applications in social media research and main studies
(differentiation between self-reported and digitally collected network data)
ethical considerations
interpretation of basic metrics

break 10 minutes

2.2. Use case: Twitter analysis - 1 hr
exploratory analysis of the selected dataset, and interpretation of results.

break 10 minutes

1 hr
3.1 Network Analysis in business research - business analytics

Overview of business analytics research - supply chain operations
business analytics research and applications of SNA - amazon dataset

Exploratory analysis of the selected dataset, and interpretation of results.



----
alternative --------------
Introduction to the analysis of socio-technical networks, and applications of network analytic techniques in LA, i.e. community detection, bipartite network analysis, network clustering, integration with text analysis. Presentation of community detection, information flow analysis, and statistical approaches in network analysis. The students will be expected to select one approach out of those presented, and implement it on one of the suggested datasets in R.
---- !!!!!!----------

2.2 Extended feautures of Graphs - 30 minutes

The Page Rank algorithm - Calculating the Google PageRank for the specified vertices.

https://www.r-bloggers.com/2019/05/benchmark-of-popular-graph-network-packages/

[Stanford Large Network Dataset Collection](https://snap.stanford.edu/data/index.html)
[amazon dataset](https://snap.stanford.edu/data/index.html#amazon) 
[google dataset](https://snap.stanford.edu/data/web-Google.html) 

break - 10 minutes

2.3. Other popular packages, benchmarking
30 minutes

graph-tool
igraph
networkit
networkx
snap

2.4. Wrap-up 



